R Packages to Aid in Handling Web Access Logs

### 2015-12-20 paper/paper.Rmd
FIRST DRAFT DONE

---
title: R Packages to Aid in Handling Web Access Logs
author:
  - name: Oliver Keyes
    affiliation: Wikimedia Foundation
    address:
    - line 1
    - line 2
    email:  author1@work
  - name: Bob Rudis
    affiliation: Rapid7
    address:
    - line 1
    - line 2
    email:  author2@work
abstract: >
  Web access logs contain information on HTTP(S) requests and form a key part of both industry and academic explorations
  of human behaviour on the internet - explorations commonly performed in R. In this paper we explain and demonstrate
  a series of packages - **webreadr**, **urltools**, **iptools** and **rgeolocate** - designed to efficiently read in, parse
  and munge access log data, allowing researchers to handle it easily.
preamble: >
output: rticles::rjournal_article
---
## Introduction

The rise of the World Wide Web over the last few decades has made it dramatically easier to access and transfer data, and R \citep{RCore} boasts abundant functionality when it comes to taking data *from* the web. Base R itself has simple file downloading and page reading capabilities, through the `download.file` and `readLines` functions, and additional functionality is made available for handling web-accessible data through packages such as **httr** \citep{httr}.

@@

As a statistically-oriented programming language, R is commonly used by these same researchers for data analysis, testing and reporting - but it lacks tools designed for the kinds of data sources and formats that are encountered when dealing with access logs. In this article we review the use cases for particular operations over web data, the limitations in base R when it comes to performing those operations, and a suite of R packages designed to overcome them: **webreadr** \citep{webreadr}, designed for reading access logs in,
**urltools** \citep{urltools}, for URL manipulation, **iptools** \citep{rgeolocate} for handling IP addresses, and **rgeolocate** \citep{rgeolocate} for direct IP geolocation.

@@

To make reading access logs into R as easy as possible we created the **webreadr** package. This contains user-friendly equivalents to `read.table` for each type of log, detecting the fields that should appear, converting the timestamps into POSIX objects, and merging fields or splitting fields where necessary. The package contains four core functions, one for each form of log: `read_clf` for the Common Log Format, `read_combined` for the Combined Log Format, and `read_squid` and `read_aws` for Squid and AWS formats respectively. Each one abstracts away the complexity of specifying column names and formats, and instead allows a researcher to read a file in with a minimal amount of work.

As the name suggests, it is built not on top of base R but on top of the **readr** \citep{readr} package, allowing us to take advantage of substantial speed improvements that package's base functions have over base R. These improvements can be seen in the visualisation below, which uses **microbenchmark** \citep{microbenchmark} to compare 100 reads of a 600,000-line "squid" formatted file with webreadr to the same operation performed in base R:

@@

To solve this common problem in analysing request logs, the **urltools** [@urltools] package was created. This contains a function,
<code>url_decode</code>, which decodes URLs - but relies on vectorised, compiled code. Benchmarking the two approaches against each other shows that the `urltools` implementation is approximately 60-70 times faster over large datasets. Again using **microbenchmark**, if we compare the vectorised decoding of 1,000,000 URLs with `urltools` against `URLdecode` in a `vapply` loop, we see a 60-70 times speed improvement of urltools over base R:

@@

Base R does not have native code to parse URLs, but the **httr** package \citep{httr} contains a function, `parse_url`, designed to do just thatm built on R's regular expressions. This function is not vectorised, does not make use of compiled code internally, and produces a list rather than data.frame, making looping over a set of URLs to parse each one a time-consuming and awkward experience. This is understandable given the intent behind that function, which is to decompose individual URLs within the context of making HTTP requests, rather than to analyse URLs *en masse*.

`urltools` contains `url_parse` - which does the same thing as the equivalent `httr` functionality, but in a vectorised way, relying on compiled code, and producing a data.frame. Within the context of parsing and processing access logs, this is far more useful, because it works efficiently over large sets: httr's functionality, not having designed to be, does not. Indeed, benchmarking showed that `url_parse` is approximately 570 times faster than httr's equivalent function:

@@

```{r, eval=TRUE}
# Set the scheme and observe the modification to the resulting URL
scheme(url) <- "https"
url
```
## IP manipulation and geolocation

## Summary

This file is only a basic article template. For full details of _The R Journal_ style and information on how to prepare your article for submission, see the [Instructions for Authors](http://journal.r-project.org/latex/RJauthorguide.pdf).

>>>>

---
title: R Packages to Aid in Handling Web Access Logs
author:
  - name: Oliver Keyes
    affiliation: Wikimedia Foundation
    address:
    - line 1
    - line 2
    email:  author1@work
  - name: Bob Rudis
    affiliation: Rapid7
    address:
    - line 1
    - line 2
    email:  author2@work
  - name: Jay Jacobs
    affiliation: Magical Python Land
    address:
    - line 1
    - line 2
    email:  author3@work
abstract: >
  Web access logs contain information on HTTP(S) requests and form a key part of both industry and academic explorations
  of human behaviour on the internet - explorations commonly performed in R. In this paper we explain and demonstrate
  a series of packages - \CRANpkg{webreadr}, \CRANpkg{urltools}, \CRANpkg{iptools} and \CRANpkg{rgeolocate} - designed to efficiently read in, parse
  and munge access log data, allowing researchers to handle it easily.
preamble: >
output: rticles::rjournal_article
---
## Introduction

The rise of the World Wide Web over the last few decades has made it dramatically easier to access and transfer data, and R \citep{RCore} boasts abundant functionality when it comes to taking data *from* the web. Base R itself has simple file downloading and page reading capabilities, through the `download.file` and `readLines` functions, and additional functionality is made available for handling web-accessible data through packages such as \CRANpkg{httr} \citep{httr}.

@@

As a statistically-oriented programming language, R is commonly used by these same researchers for data analysis, testing and reporting - but it lacks tools designed for the kinds of data sources and formats that are encountered when dealing with access logs. In this article we review the use cases for particular operations over web data, the limitations in base R when it comes to performing those operations, and a suite of R packages designed to overcome them: \CRANpkg{webreadr} \citep{webreadr}, designed for reading access logs in,
\CRANpkg{urltools} \citep{urltools}, for URL manipulation, \CRANpkg{iptools} \citep{iptools} for handling IP addresses, and \CRANpkg{rgeolocate} \citep{rgeolocate} for direct IP geolocation.

@@

To make reading access logs into R as easy as possible we created the \CRANpkg{webreadr} package. This contains user-friendly equivalents to `read.table` for each type of log, detecting the fields that should appear, converting the timestamps into POSIX objects, and merging fields or splitting fields where necessary. The package contains four core functions, one for each form of log: `read_clf` for the Common Log Format, `read_combined` for the Combined Log Format, and `read_squid` and `read_aws` for Squid and AWS formats respectively. Each one abstracts away the complexity of specifying column names and formats, and instead allows a researcher to read a file in with a minimal amount of work.

As the name suggests, it is built not on top of base R but on top of the \CRANpkg{readr} \citep{readr} package, allowing us to take advantage of substantial speed improvements that package's base functions have over base R. These improvements can be seen in the visualisation below, which uses \CRANpkg{microbenchmark} \citep{microbenchmark} to compare 100 reads of a 600,000-line "squid" formatted file with webreadr to the same operation performed in base R:

@@

To solve this common problem in analysing request logs, the \CRANpkg{urltools} [@urltools] package was created. This contains a function,
<code>url_decode</code>, which decodes URLs - but relies on vectorised, compiled code. Benchmarking the two approaches against each other shows that the \CRANpkg{urltools} implementation is approximately 60-70 times faster over large datasets. Again using **microbenchmark**, if we compare the vectorised decoding of 1,000,000 URLs with \CRANpkg{urltools} against `URLdecode` in a `vapply` loop, we see a 60-70 times speed improvement of urltools over base R:

@@

Base R does not have native code to parse URLs, but the \CRANpkg{httr} package \citep{httr} contains a function, `parse_url`, designed to do just thatm built on R's regular expressions. This function is not vectorised, does not make use of compiled code internally, and produces a list rather than data.frame, making looping over a set of URLs to parse each one a time-consuming and awkward experience. This is understandable given the intent behind that function, which is to decompose individual URLs within the context of making HTTP requests, rather than to analyse URLs *en masse*.

\CRANpkg{urltools} contains `url_parse` - which does the same thing as the equivalent `httr` functionality, but in a vectorised way, relying on compiled code, and producing a data.frame. Within the context of parsing and processing access logs, this is far more useful, because it works efficiently over large sets: httr's functionality, which was never designed with vectorisation in mind, does not. Indeed, benchmarking showed that `url_parse` is approximately 570 times faster than httr's equivalent function:

@@

```{r, eval=TRUE}
# Set the scheme and observe the modification to the resulting URL
scheme(url) <- "https"
url
```

As a result of this functionality, urltools provides a package to make URL manipulation faster, easier and far more accessible, reducing the burden associated with munging access logs or similar datasets and allowing a researcher to get to the statistical analysis faster.

## IP manipulation and geolocation

Access logs also contain IP addresses - unique numeric values that identify a particular computer or network in the context of the internet.

<!-- Bob, I'll rely on you to fill out bits on how iptools is useful for this kind of work; I tend to primarily rely on rgeolocate since my ip-needs are more limited. -->

As a side-effect of how IP addresses tend to be assigned - in geographic blocks, to individual machines or to local networks - they can be used to geolocate requests, identifying where in the world the request came from, sometimes down to the level of individual post codes or pairs of latitude/longitude coordinates.

This is tremendously useful in industry, where the geographic reach of a service has substantial implications for its viability and survivability, and in academia, where the locality of internet-provided information and the breadth of internet access are active concerns and areas of study \citep{barriers}.

Many services and databases exist for extracting geographic metadata from IP addresses. One of the most common is the service provided by MaxMind, which has both proprietary and openly-licensed databases, in binary and comma-separated formats. The free databases have been used by various web APIs, which makes the data they contain accessible from R. Unfortunately, dependence on web APIs means that handling large numbers of IP addresses can be very slow (they tend to be designed to only accept one IP address at a time, and may contain throttling beyond that) and has privacy concerns, since it essentially means sending user IP addresses to a third party. And even without these issues, there are no convenient wrappers for those APIs available on CRAN: users have to hand-roll their own.

With these concerns in mind, we wrote the \CRANpkg{rgeolocate} package. Through \pkg{httr} this contains convenient, vectorised bindings
to various web services that provide geographic metadata about IP addresses. More importantly, using the \CRANpkg{Rcpp} package to integrate C++ and R, \pkg{rgeolocate} also features a direct, compiled binding to the MaxMind API. This means that local binary databases can also be queried, which is far faster and more robust than web-based equivalents and avoids the privacy concerns associated with transmitting users' IP addresses externally.

The MaxMind API requires a paid or free binary database - one of which, for country-level IP resolution, is included in \pkg{rgeolocate} - and allows you to retrieve the continent, country name or ISO code, region or city name, tzdata-compatible timezone, longitude, latitude or connection type of a particular IP address. Multiple fields can be selected (although which are available depends on the type of database used), and results are returned in a data.frame:

```{r, eval=TRUE}

# Load rgeolocate
library(rgeolocate)

# Find the rgeolocate-provided binary database
geo_file <- system.file("extdata","GeoLite2-Country.mmdb", package = "rgeolocate")

# Put together some example IP addresses
ip_addresses <- c("174.62.175.82", "196.200.60.51")

# Geolocate
rgeolocate::maxmind(ip_addresses, geo_file, fields = c("continent_name", "country_code"))

```

Direct speed comparisons aren't possible, since the functionality it provides is not (to the authors' knowledge) replicated in
other R packages, but it is certainly faster (and more secure) than internet-dependent alternatives.

## Conclusions and further work

In this research article we have demonstrated a set of tools for handling access logs during every stage of the data cleaning pipeline - reading them in with \pkg{webreadr}, decoding, manipulating and extracting value from URLs with \pkg{urltools}, and retrieving geographic metadata from IP addresses with \pkg{iptools} and \pkg{rgeolocate}. We have also demonstrated the dramatic speed improvements in using these tools in preference to existing methods within R.

Further work - integrating more sources of geolocation information within \pkg{rgeolocate}, supporting UTF-8 URL decoding within \pkg{urltools}, and increasing IPv6 support in \pkg{iptools} - would make these packages even more useful to Human-Computer Interaction researchers and other specialists who rely on access logs as primary data sources. Even without that functionality, however, this suite of packages is a dramatic improvement upon the status quo.

### 2015-12-20 paper/paper.tex
FIRST DRAFT DONE

% !TeX root = RJwrapper.tex
\title{R Packages to Aid in Handling Web Access Logs}
\author{by Oliver Keyes, Bob Rudis}

@@

\abstract{
Web access logs contain information on HTTP(S) requests and form a key
part of both industry and academic explorations of human behaviour on
the internet - explorations commonly performed in R. In this paper we
explain and demonstrate a series of packages - \textbf{webreadr},
\textbf{urltools}, \textbf{iptools} and \textbf{rgeolocate} - designed
to efficiently read in, parse and munge access log data, allowing
researchers to handle it easily.
}

@@

The rise of the World Wide Web over the last few decades has made it
dramatically easier to access and transfer data, and R \citep{RCore}
boasts abundant functionality when it comes to taking data \emph{from}
the web. Base R itself has simple file downloading and page reading
capabilities, through the \texttt{download.file} and \texttt{readLines}
functions, and additional functionality is made available for handling
web-accessible data through packages such as \textbf{httr} \citep{httr}.

@@

As a statistically-oriented programming language, R is commonly used by
these same researchers for data analysis, testing and reporting - but it
lacks tools designed for the kinds of data sources and formats that are
encountered when dealing with access logs. In this article we review the
use cases for particular operations over web data, the limitations in
base R when it comes to performing those operations, and a suite of R
packages designed to overcome them: \textbf{webreadr} \citep{webreadr},
designed for reading access logs in, \textbf{urltools} \citep{urltools},
for URL manipulation, \textbf{iptools} \citep{rgeolocate} for handling
IP addresses, and \textbf{rgeolocate} \citep{rgeolocate} for direct IP
geolocation.

@@

To make reading access logs into R as easy as possible we created the
\textbf{webreadr} package. This contains user-friendly equivalents to
\texttt{read.table} for each type of log, detecting the fields that
should appear, converting the timestamps into POSIX objects, and merging
fields or splitting fields where necessary. The package contains four
core functions, one for each form of log: \texttt{read\_clf} for the
Common Log Format, \texttt{read\_combined} for the Combined Log Format,
and \texttt{read\_squid} and \texttt{read\_aws} for Squid and AWS
formats respectively. Each one abstracts away the complexity of
specifying column names and formats, and instead allows a researcher to
read a file in with a minimal amount of work.

As the name suggests, it is built not on top of base R but on top of the
\textbf{readr} \citep{readr} package, allowing us to take advantage of
substantial speed improvements that package's base functions have over
base R. These improvements can be seen in the visualisation below, which
uses \textbf{microbenchmark} \citep{microbenchmark} to compare 100 reads
of a 600,000-line ``squid'' formatted file with webreadr to the same
operation performed in base R:

@@

To solve this common problem in analysing request logs, the
\textbf{urltools} {[}@urltools{]} package was created. This contains a
function, url\_decode, which decodes URLs - but relies on vectorised,
compiled code. Benchmarking the two approaches against each other shows
that the \texttt{urltools} implementation is approximately 60-70 times
faster over large datasets. Again using \textbf{microbenchmark}, if we
compare the vectorised decoding of 1,000,000 URLs with \texttt{urltools}
against \texttt{URLdecode} in a \texttt{vapply} loop, we see a 60-70
times speed improvement of urltools over base R:

@@

Base R does not have native code to parse URLs, but the \textbf{httr}
package \citep{httr} contains a function, \texttt{parse\_url}, designed
to do just thatm built on R's regular expressions. This function is not
vectorised, does not make use of compiled code internally, and produces
a list rather than data.frame, making looping over a set of URLs to
parse each one a time-consuming and awkward experience. This is
understandable given the intent behind that function, which is to
decompose individual URLs within the context of making HTTP requests,
rather than to analyse URLs \emph{en masse}.

\texttt{urltools} contains \texttt{url\_parse} - which does the same
thing as the equivalent \texttt{httr} functionality, but in a vectorised
way, relying on compiled code, and producing a data.frame. Within the
context of parsing and processing access logs, this is far more useful,
because it works efficiently over large sets: httr's functionality, not
having designed to be, does not. Indeed, benchmarking showed that
\texttt{url\_parse} is approximately 570 times faster than httr's
equivalent function:

@@

A vector of URLs passed into \texttt{url\_parse} produces a data.frame,
with one column for each of the IETF-supported components, and empty
strings representing components that could not be found. Additionally,
influenced by the style of the \emph{lubridate} package
\citep{lubridate}, \texttt{urltools} contains functions to get or set
\emph{individual} components:

\begin{Schunk}
\begin{Sinput}
library(urltools)
url <- "http://www.google.com/"
scheme(url)
\end{Sinput}
\begin{Soutput}
#> [1] "http"
\end{Soutput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
scheme(url) <- "https"
url
\end{Sinput}
\begin{Soutput}
#> [1] "https://www.google.com/"
\end{Soutput}
\end{Schunk}

@@

@@

>>>>

% !TeX root = RJwrapper.tex
\title{R Packages to Aid in Handling Web Access Logs}
\author{by Oliver Keyes, Bob Rudis, Jay Jacobs}

@@

\abstract{
Web access logs contain information on HTTP(S) requests and form a key
part of both industry and academic explorations of human behaviour on
the internet - explorations commonly performed in R. In this paper we
explain and demonstrate a series of packages - \CRANpkg{webreadr},
\CRANpkg{urltools}, \CRANpkg{iptools} and \CRANpkg{rgeolocate} -
designed to efficiently read in, parse and munge access log data,
allowing researchers to handle it easily.
}

@@

The rise of the World Wide Web over the last few decades has made it
dramatically easier to access and transfer data, and R \citep{RCore}
boasts abundant functionality when it comes to taking data \emph{from}
the web. Base R itself has simple file downloading and page reading
capabilities, through the \texttt{download.file} and \texttt{readLines}
functions, and additional functionality is made available for handling
web-accessible data through packages such as \CRANpkg{httr}
\citep{httr}.

@@

As a statistically-oriented programming language, R is commonly used by
these same researchers for data analysis, testing and reporting - but it
lacks tools designed for the kinds of data sources and formats that are
encountered when dealing with access logs. In this article we review the
use cases for particular operations over web data, the limitations in
base R when it comes to performing those operations, and a suite of R
packages designed to overcome them: \CRANpkg{webreadr} \citep{webreadr},
designed for reading access logs in, \CRANpkg{urltools}
\citep{urltools}, for URL manipulation, \CRANpkg{iptools}
\citep{iptools} for handling IP addresses, and \CRANpkg{rgeolocate}
\citep{rgeolocate} for direct IP geolocation.

@@

To make reading access logs into R as easy as possible we created the
\CRANpkg{webreadr} package. This contains user-friendly equivalents to
\texttt{read.table} for each type of log, detecting the fields that
should appear, converting the timestamps into POSIX objects, and merging
fields or splitting fields where necessary. The package contains four
core functions, one for each form of log: \texttt{read\_clf} for the
Common Log Format, \texttt{read\_combined} for the Combined Log Format,
and \texttt{read\_squid} and \texttt{read\_aws} for Squid and AWS
formats respectively. Each one abstracts away the complexity of
specifying column names and formats, and instead allows a researcher to
read a file in with a minimal amount of work.

As the name suggests, it is built not on top of base R but on top of the
\CRANpkg{readr} \citep{readr} package, allowing us to take advantage of
substantial speed improvements that package's base functions have over
base R. These improvements can be seen in the visualisation below, which
uses \CRANpkg{microbenchmark} \citep{microbenchmark} to compare 100
reads of a 600,000-line ``squid'' formatted file with webreadr to the
same operation performed in base R:

@@

To solve this common problem in analysing request logs, the
\CRANpkg{urltools} {[}@urltools{]} package was created. This contains a
function, url\_decode, which decodes URLs - but relies on vectorised,
compiled code. Benchmarking the two approaches against each other shows
that the \CRANpkg{urltools} implementation is approximately 60-70 times
faster over large datasets. Again using \textbf{microbenchmark}, if we
compare the vectorised decoding of 1,000,000 URLs with
\CRANpkg{urltools} against \texttt{URLdecode} in a \texttt{vapply} loop,
we see a 60-70 times speed improvement of urltools over base R:

@@

Base R does not have native code to parse URLs, but the \CRANpkg{httr}
package \citep{httr} contains a function, \texttt{parse\_url}, designed
to do just thatm built on R's regular expressions. This function is not
vectorised, does not make use of compiled code internally, and produces
a list rather than data.frame, making looping over a set of URLs to
parse each one a time-consuming and awkward experience. This is
understandable given the intent behind that function, which is to
decompose individual URLs within the context of making HTTP requests,
rather than to analyse URLs \emph{en masse}.

\CRANpkg{urltools} contains \texttt{url\_parse} - which does the same
thing as the equivalent \texttt{httr} functionality, but in a vectorised
way, relying on compiled code, and producing a data.frame. Within the
context of parsing and processing access logs, this is far more useful,
because it works efficiently over large sets: httr's functionality,
which was never designed with vectorisation in mind, does not. Indeed,
benchmarking showed that \texttt{url\_parse} is approximately 570 times
faster than httr's equivalent function:

@@

A vector of URLs passed into \texttt{url\_parse} produces a data.frame,
with one column for each of the IETF-supported components, and empty
strings representing components that could not be found. Additionally,
influenced by the style of the \emph{lubridate} package
\citep{lubridate}, \texttt{urltools} contains functions to get or
\emph{set} individual components:

\begin{Schunk}
\begin{Sinput}
# Load urltools and construct a URL
library(urltools)
url <- "http://www.google.com/"

# Get the scheme
scheme(url)
\end{Sinput}
\begin{Soutput}
#> [1] "http"
\end{Soutput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
# Set the scheme and observe the modification to the resulting URL
scheme(url) <- "https"
url
\end{Sinput}
\begin{Soutput}
#> [1] "https://www.google.com/"
\end{Soutput}
\end{Schunk}

As a result of this functionality, urltools provides a package to make
URL manipulation faster, easier and far more accessible, reducing the
burden associated with munging access logs or similar datasets and
allowing a researcher to get to the statistical analysis faster.

@@

Access logs also contain IP addresses - unique numeric values that
identify a particular computer or network in the context of the
internet.

As a side-effect of how IP addresses tend to be assigned - in geographic
blocks, to individual machines or to local networks - they can be used
to geolocate requests, identifying where in the world the request came
from, sometimes down to the level of individual post codes or pairs of
latitude/longitude coordinates.

This is tremendously useful in industry, where the geographic reach of a
service has substantial implications for its viability and
survivability, and in academia, where the locality of internet-provided
information and the breadth of internet access are active concerns and
areas of study \citep{barriers}.

Many services and databases exist for extracting geographic metadata
from IP addresses. One of the most common is the service provided by
MaxMind, which has both proprietary and openly-licensed databases, in
binary and comma-separated formats. The free databases have been used by
various web APIs, which makes the data they contain accessible from R.
Unfortunately, dependence on web APIs means that handling large numbers
of IP addresses can be very slow (they tend to be designed to only
accept one IP address at a time, and may contain throttling beyond that)
and has privacy concerns, since it essentially means sending user IP
addresses to a third party. And even without these issues, there are no
convenient wrappers for those APIs available on CRAN: users have to
hand-roll their own.

With these concerns in mind, we wrote the \CRANpkg{rgeolocate} package.
Through \pkg{httr} this contains convenient, vectorised bindings to
various web services that provide geographic metadata about IP
addresses. More importantly, using the \CRANpkg{Rcpp} package to
integrate C++ and R, \pkg{rgeolocate} also features a direct, compiled
binding to the MaxMind API. This means that local binary databases can
also be queried, which is far faster and more robust than web-based
equivalents and avoids the privacy concerns associated with transmitting
users' IP addresses externally.

The MaxMind API requires a paid or free binary database - one of which,
for country-level IP resolution, is included in \pkg{rgeolocate} - and
allows you to retrieve the continent, country name or ISO code, region
or city name, tzdata-compatible timezone, longitude, latitude or
connection type of a particular IP address. Multiple fields can be
selected (although which are available depends on the type of database
used), and results are returned in a data.frame:

\begin{Schunk}
\begin{Sinput}
# Load rgeolocate
library(rgeolocate)

# Find the rgeolocate-provided binary database
geo_file <- system.file("extdata","GeoLite2-Country.mmdb", package = "rgeolocate")

# Put together some example IP addresses
ip_addresses <- c("174.62.175.82", "196.200.60.51")

# Geolocate
rgeolocate::maxmind(ip_addresses, geo_file, fields = c("continent_name", "country_code"))
\end{Sinput}
\begin{Soutput}
#>   continent_name country_code
#> 1  North America           US
#> 2         Africa           ML
\end{Soutput}
\end{Schunk}

@@

\address{
Jay Jacobs\\
Magical Python Land\\
line 1\\ line 2\\
}
\href{mailto:author3@work}{\nolinkurl{author3@work}}
